|

### 2024-11-26 manuscript/appendix2.Rmd
renamed files to remove chapter2 in title
files up to date
arranged appendix2 rmd
update all appendix files

---
title: |
     Experience shapes individual foraging specialization and success in a virtual predator-prey system: \
     Appendix 2

output: 
  word_document:
    reference_docx: Chapter2-styles-appendix.docx
    fig_caption: yes
csl: ecology.csl
---

@@

```{r tableprep, include = FALSE}

# ===========================================================================
# 1. Extract the fixed effects from the MV model
# ===========================================================================

@@

# Model fit
fit <- readRDS(
  here::here("outputs", "01_outputs_models", "B2_DHMLM.rds")
)

# Fixed effects table
results <- data.table(
  round(summary(fit)$fixed, digits = 2)[, 1:4],
  keep.rownames = TRUE
)

setnames(results, "rn", "parameter")

results[parameter %like% "predspeed", trait := "predator speed"][
  parameter %like% "preyavgspeed", trait := "prey speed"][
    parameter %like% "success", trait := "hunting success"
  ]

# mu parameters table
tabs1 <- results[
    !(parameter %like% "sigma"),
    c(6, 1, 2, 4, 5)
]

# sigma parameters table
tabs2 <- results[
    parameter %like% "sigma",
    c(6, 1, 2, 4, 5)
]

# Rename parameter values ---------------------------------------------------

# Rename
tabs1[,
  parameter := c(
    rep(
      c("prey rank",
        "cumulative experience",
        "group 1",
        "group 4",
        "group 2",
        "group 3"),
      2
    ),
      c("game duration",
        "cumulative experience",
        "group 1",
        "group 4",
        "group 2",
        "group 3")
  )
]
# Reorder rows
tabs1 <- tabs1[
  c(1:3, 5, 6, 4,
    7:9, 11, 12, 10,
    13:15, 17, 18, 16)
]

# Rename
tabs2[,
  parameter := rep(
    c("prey rank",
      "cumulative experience",
      "group 1",
      "group 4",
      "group 2",
      "group 3"),
    2
  )
]

# Reorder
tabs2 <- tabs2[
  c(1:3, 5, 6, 4,
    7:9, 11, 12, 10)
]

@@

# Custom header
my_header <- data.frame(
   col_keys = c("trait",
                "parameter",
                "Estimate",
                "l-95% CI",
                "u-95% CI"),
   line1 = c("Trait",
             "Parameter",
             "Value",
             "lower 95% CI",
             "upper 95% CI"),
   stringsAsFactors = FALSE
)

# Custom theme
my_theme <- function(x, ...) {
   x <- colformat_double(x, big.mark = " ",
                         decimal.mark = ".",
                         digits = 2)
   x <- set_table_properties(x, layout = "fixed")
   x <- border_remove(x)
   std_border <- fp_border(width = 1, color = "black")
   x <- hline_top(x, part = "all", border = std_border)
   x <- hline_bottom(x, part = "all", border = std_border)
   autofit(x)
}

```{r tableS1, ft.align = "left"}
# Create the table ----------------------------------------------------------

tabs1f <- tabs1 %>%
  flextable(col_keys = my_header$col_keys) %>%
  set_header_df(mapping = my_header, key = "col_keys") %>%
  my_theme() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top", part = "body") %>%
  align(align = "left", part = "all", j = c(1, 2)) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5) %>%
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "Group 1: <50 matches, Group 2: between 50 and 99 matches, Group 3: between 100 and 299 matches, Group 4: > 299 (i.e. group in the main text)"
    ),
    ref_symbols = " "
  ) %>%
  set_caption(
    tabs1,
    caption = "Table S1. Fixed effects table on the mean part of the DHMLM comparing random groups of players with different matches played to the group presented in the main text.",
    style = "Table Caption"
  ) %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(size = 12, part = "footer") %>%
  font(fontname = "Times New Roman", part = "all")

tabs1f

@@

tabs2f <- tabs2 %>%
  flextable(col_keys = my_header$col_keys) %>%
  set_header_df(mapping = my_header, key = "col_keys") %>%
  my_theme() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top", part = "body") %>%
  align(align = "left", part = "all", j = c(1, 2)) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5) %>%
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "Group 1: <50 matches, Group 2: between 50 and 99 matches, Group 3: between 100 and 299 matches, Group 4: > 299 (i.e. group in the main text)"
    ),
    ref_symbols = " "
  ) %>%
  set_caption(
    tabs2,
    caption = "Table S2. Fixed effects table on the dispersion part of the DHMLM comparing random groups of players with different matches played to the group presented in the main text.",
    style = "Table Caption"
  ) %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(size = 12, part = "footer") %>%
  font(fontname = "Times New Roman", part = "all")

tabs2f

@@

>>>>

---
title: |
     Individual foraging specialization and success change across experience in a virtual predator-prey system: \
     Appendix 3

output: 
  word_document:
    reference_docx: appendix-styles.docx
    fig_caption: yes
csl: ecology.csl
---

@@

```{r tableS1, ft.align = "left"}

# ===========================================================================
# 1. Import the model fit
# ===========================================================================

# path
path <- file.path(here::here("outputs"))

# Import the model
fit <- readRDS(
  file.path(path, "01_outputs_models", "B1_DHMLM.rds")
)

# ===========================================================================
# 2. Fixed effects table
# ===========================================================================

@@

# Extract the fixed effects posterior samples
b_draws <- data.table(
  as_draws_df(
    fit, variable = "^b_", regex = TRUE
  )
)

b_draws[, c(34:36) := NULL]

# Back-transform the parameters ----------------------------------------------

# Take the exponent of sigma parameters
b_draws[
  , c("b_sigma_speednovice_Intercept",
      "b_sigma_speedinterm_Intercept",
      "b_sigma_speedadvanced_Intercept",
      "b_sigma_preyspeednovice_Intercept",
      "b_sigma_preyspeedinterm_Intercept",
      "b_sigma_preyspeedadvanced_Intercept",
      "b_sigma_speednovice_Zprey_avg_rank",
      "b_sigma_speedinterm_Zprey_avg_rank",
      "b_sigma_speedadvanced_Zprey_avg_rank",
      "b_sigma_preyspeednovice_Zprey_avg_rank",
      "b_sigma_preyspeedinterm_Zprey_avg_rank",
      "b_sigma_preyspeedadvanced_Zprey_avg_rank"
  )
  := lapply(.SD, function(x) {exp(x)}),
  .SDcols = c(
    "b_sigma_speednovice_Intercept",
    "b_sigma_speedinterm_Intercept",
    "b_sigma_speedadvanced_Intercept",
    "b_sigma_preyspeednovice_Intercept",
    "b_sigma_preyspeedinterm_Intercept",
    "b_sigma_preyspeedadvanced_Intercept",
    "b_sigma_speednovice_Zprey_avg_rank",
    "b_sigma_speedinterm_Zprey_avg_rank",
    "b_sigma_speedadvanced_Zprey_avg_rank",
    "b_sigma_preyspeednovice_Zprey_avg_rank",
    "b_sigma_preyspeedinterm_Zprey_avg_rank",
    "b_sigma_preyspeedadvanced_Zprey_avg_rank"
  )
]

# plogis for the success parameters
b_draws[
  , c("b_successnovice_Intercept",
      "b_successinterm_Intercept",
      "b_successadvanced_Intercept",
      "b_successnovice_Zgame_duration",
      "b_successinterm_Zgame_duration",
      "b_successadvanced_Zgame_duration",
      "b_successnovice_Zprey_avg_rank",
      "b_successinterm_Zprey_avg_rank",
      "b_successadvanced_Zprey_avg_rank"
  )
  := lapply(.SD, function(x) {plogis(x)}),
  .SDcols = c(
    "b_successnovice_Intercept",
    "b_successinterm_Intercept",
    "b_successadvanced_Intercept",
    "b_successnovice_Zgame_duration",
    "b_successinterm_Zgame_duration",
    "b_successadvanced_Zgame_duration",
    "b_successnovice_Zprey_avg_rank",
    "b_successinterm_Zprey_avg_rank",
    "b_successadvanced_Zprey_avg_rank"
  )
]

# Reshape the table -----------------------------------------------------------

fixef_tab <- melt(
  b_draws,
  variable.name = "parameter"
)

# Summarize the values (medians + 95% CI) -------------------------------------

# Intervals
lower_interval <- function(x) {
  coda::HPDinterval(as.mcmc(x), 0.95)[1]
}
upper_interval <- function(x) {
  coda::HPDinterval(as.mcmc(x), 0.95)[2]
}

# Summarize values
fixef_tab[, ":=" (median = median(value),
                  lower_ci = lower_interval(value),
                  upper_ci = upper_interval(value)),
          by = .(parameter)]

fixef_tab <- unique(fixef_tab[, .(parameter, median, lower_ci, upper_ci)])
fixef_tab[, c(2:4) := round(fixef_tab[, c(2:4)], digits = 2)]

# Rearrange the table --------------------------------------------------------

# Add a column that specifies the experience level
fixef_tab[parameter %like% "novice", xp_level := "novice"]
fixef_tab[parameter %like% "interm", xp_level := "intermediate"]
fixef_tab[parameter %like% "advanced", xp_level := "advanced"]

# Paste upper and lower ci with estimate
fixef_tab[, estimate := paste(format(median, digits = 2), "(")]
fixef_tab[, estimate := paste(estimate, format(lower_ci, digits = 2), sep = "")]
fixef_tab[, estimate := paste(estimate, ",", sep = "")]
fixef_tab[, estimate := paste(estimate, format(upper_ci, digits = 2), sep = " ")]
fixef_tab[, estimate := paste(estimate, ")", sep = "")]
fixef_tab[, c("median", "lower_ci", "upper_ci") := NULL]

# reorder columns, and then rows based on xp level
setorder(
  fixef_tab,
  cols = -"xp_level"
)

# Create trait column
fixef_tab[, trait := ""][
  parameter %like% "success", trait := "hunting success"][
    parameter %like% "speed", trait := "predator speed"][
      parameter %like% "preyspeed", trait := "prey speed"
      ]

# Rename parameter column
fixef_tab[
  , parameter := rep(
    c(
      rep(c("intercept (mean)", "intercept (sigma)"), 2),
      "intercept (mean)",
      rep(c("prey rank (mean)", "prey rank (sigma)"), 2),
      "match duration (mean)", "prey rank (mean)"
    ),
    3
  )
]

# Reshape
fixef_tab <- dcast(
  fixef_tab,
  trait + parameter ~ xp_level,
  value.var = "estimate"
)

# Reorder columns and rows
fixef_tab <- fixef_tab[, c(1, 2, 5, 4, 3)]
table <- fixef_tab[c(4, 6, 5, 7, 8, 10, 9, 11, 1, 2, 3)]

# Because I can't align the rows properly
# in the flextable object
table[c(2:4, 6:8, 10, 11), 1] <- " "

@@

# Custom header
my_header <- data.frame(
  col_keys = c("trait", "parameter",
               "novice", "intermediate", "advanced"),
  line1 = c("Trait",
            "Parameter",
            "Novice",
            "Intermediate",
            "Advanced"),
  stringsAsFactors = FALSE
)

# Custom theme
my_theme <- function(x, ...) {
  x <- colformat_double(x, big.mark = " ",
                        decimal.mark = ".",
                        digits = 2)
  x <- border_remove(x)
  std_border <- fp_border(width = 1, color = "black")
  x <- hline_top(x, part = "all", border = std_border)
  x <- hline_bottom(x, part = "all", border = std_border)
  x <- width(x, j = 1, width = 2.84, unit = "cm")
  x <- width(x, j = 2, width = 3.78, unit = "cm")
  x <- width(x, j = c(3:5), width = 3.3, unit = "cm")
}

# Create the table ----------------------------------------------------------

tab1 <-

  # table structure
  table %>%
  select(trait, parameter,
         novice, intermediate, advanced) %>%
  flextable(col_keys = my_header$col_keys) %>%
  set_header_df(mapping = my_header, key = "col_keys") %>%
  my_theme() %>%

  # Align repeated rows on column 1
  #merge_v(j = 1) %>%
  valign(j = 1, valign = "top", part = "body") %>%

  # Align text
  align(align = "left", part = "all", j = 1) %>%
  align(align = "left", part = "all", j = 2) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5) %>%

  # Footnote
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "We exponentiated the dispersion parameters (i.e. sigma) which are estimated on a log scale. We back-transformed the hunting success values, estimated on a logit scale, back to a probability scale.",
      as_sup("\nb "),
      "The intercept values on the mean part of the equation for all traits indicate mean behaviour and success at the population level. The intercept values on the dispersion (i.e. sigma) part of the equation for predator speed indicate behavioural specialization at the population level."
    ),
    ref_symbols = " "
  ) %>%

  # Table caption
  set_caption(
    caption = "Table S1. Posterior medians and 95% HPD intervals of the fixed effects estimated by the MDHGLM of predator speed, prey speed, and predator hunting success. The coefficients are from the model that includes the outlier.",
    style = "Table Caption"
  ) %>%

  # Font options
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(size = 12, part = "footer")

tab1

@@

# ===========================================================================
# 1. Import the model fit
# ===========================================================================

# path
path <- file.path(here::here("outputs"))

# Import the model
fit <- readRDS(
  file.path(path, "01_outputs_models", "B1_DHMLM.rds")
)

# ===========================================================================
# 2. Random effects table
# ===========================================================================

# Prepare data --------------------------------------------------------------

# Extract the random effect variances
sd_draws <- data.table(
   as_draws_df(
      fit, variable = "^sd_", regex = TRUE
   )
)

sd_draws[, c(28:30) := NULL]

# Calculate parameters -------------------------------------------------------

# Take the exponent of sigma parameters
sd_draws[
   , c("sd_predator_id__sigma_speednovice_Intercept",
       "sd_predator_id__sigma_speedinterm_Intercept",
       "sd_predator_id__sigma_speedadvanced_Intercept",
       "sd_predator_id__sigma_preyspeednovice_Intercept",
       "sd_predator_id__sigma_preyspeedinterm_Intercept",
       "sd_predator_id__sigma_preyspeedadvanced_Intercept"
   )
   := lapply(.SD, function(x) {exp(x)}),
   .SDcols = c(
      "sd_predator_id__sigma_speednovice_Intercept",
      "sd_predator_id__sigma_speedinterm_Intercept",
      "sd_predator_id__sigma_speedadvanced_Intercept",
      "sd_predator_id__sigma_preyspeednovice_Intercept",
      "sd_predator_id__sigma_preyspeedinterm_Intercept",
      "sd_predator_id__sigma_preyspeedadvanced_Intercept"
   )
]

# Reshape the table -----------------------------------------------------------

ranef_tab <- melt(
   sd_draws,
   variable.name = "parameter"
)

# Summarize the values (means + 95% CI) --------------------------------------

# Intervals
lower_interval <- function(x) {
  coda::HPDinterval(as.mcmc(x), 0.95)[1]
}
upper_interval <- function(x) {
  coda::HPDinterval(as.mcmc(x), 0.95)[2]
}

# Summarize values
ranef_tab[, ":=" (median = median(value),
                  lower_ci = lower_interval(value),
                  upper_ci = upper_interval(value)),
          by = .(parameter)]

ranef_tab <- unique(ranef_tab[, .(parameter, median, lower_ci, upper_ci)])
ranef_tab[, c(2:4) := round(ranef_tab[, c(2:4)], digits = 2)]

# Rearrange the table --------------------------------------------------------

# Add a column that specifies the experience level
ranef_tab[parameter %like% "novice", xp_level := "novice"]
ranef_tab[parameter %like% "interm", xp_level := "intermediate"]
ranef_tab[parameter %like% "advanced", xp_level := "advanced"]

# Paste upper and lower ci with estimate
ranef_tab[, estimate := paste(format(median, digits = 2), "(")]
ranef_tab[, estimate := paste(estimate, format(lower_ci, digits = 2), sep = "")]
ranef_tab[, estimate := paste(estimate, ",", sep = "")]
ranef_tab[, estimate := paste(estimate, format(upper_ci, digits = 2), sep = " ")]
ranef_tab[, estimate := paste(estimate, ")", sep = "")]
ranef_tab[, c("median", "lower_ci", "upper_ci") := NULL]

# Parameter column
ranef_tab[
   , parameter1 := ifelse(
       parameter %like% "predator_id__sigma",
       "predator ID (sigma)",
       "predator ID (mean)"
   )
]
ranef_tab[parameter %like% "avatar", parameter1 := "avatar (mean)"]
ranef_tab[parameter %like% "env", parameter1 := "environment (mean)"]

# Trait column
ranef_tab[
   , trait := ifelse(
       parameter %like% "preyspeed",
       "prey speed",
       "predator speed"
   )
]
ranef_tab[parameter %like% "success", trait := "hunting success"]

# Remove old parameter column and replace by new
ranef_tab[, parameter := NULL]
setnames(ranef_tab, "parameter1", "parameter")

# Reshape
ranef_tab <- dcast(
   ranef_tab,
   trait + parameter ~ xp_level,
   value.var = "estimate"
)

# Reorder columns
ranef_tab <- ranef_tab[, c(1, 2, 5, 4, 3)]

# Reorder rows
table <- ranef_tab[c(2:9, 1)]

# ===========================================================================
# 3. Create the table using flextable
# ===========================================================================

# Custom header
my_header <- data.frame(
  col_keys = c("trait", "parameter",
               "novice", "intermediate", "advanced"),
  line1 = c("Trait",
            "Parameter",
            "Novice",
            "Intermediate",
            "Advanced"),
  stringsAsFactors = FALSE
)

# Custom theme
my_theme <- function(x, ...) {
  x <- colformat_double(x, big.mark = " ",
                        decimal.mark = ".",
                        digits = 2)
  x <- set_table_properties(x, layout = "autofit")
  x <- border_remove(x)
  std_border <- fp_border(width = 1, color = "black")
  x <- hline_top(x, part = "all", border = std_border)
  x <- hline_bottom(x, part = "all", border = std_border)
}

# Create the table ----------------------------------------------------------

tab2 <-

  # table structure
  table %>%
  select(trait, parameter,
         novice, intermediate, advanced) %>%
  flextable(col_keys = my_header$col_keys) %>%
  set_header_df(mapping = my_header, key = "col_keys") %>%
  my_theme() %>%

  # Align repeated rows on column 1
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top", part = "body") %>%
  # font
  fontsize(size = 10, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%

  # Align text
  align(align = "left", part = "all", j = 1) %>%
  align(align = "left", part = "all", j = 2) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5) %>%

  # Add footnote
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "We exponentiated the dispersion parameters (i.e. sigma) which are estimated on a log scale.",
      as_sup("\nb "),
      "The standard deviation values on the mean part of the equation indicate, for all traits, among individual differences in mean behaviour, prey encountered, and success.",
      as_sup("\nc "),
      "The standard deviation values on the dispersion part of the equation (i.e. sigma) for predator speed indicate among individual differences in behavioural specialization. For prey speed, they indicate among individual differences in the variability of prey encounters."
    ),
    ref_symbols = " "
  ) %>%

  # Add table caption
  set_caption(
    caption = "Table S2. Posterior medians and 95% HPD intervals of the random effect standard deviations estimated by the MDHGLM of predator speed, prey speed, and predator hunting success. The coefficients are from the model that includes the outlier.",
    style = "Table Caption"
  ) %>%

  # Font options
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(size = 12, part = "footer")

tab2

@@

```{r figS1, out.width = "100%", fig.cap = "**Figure S1.** Posterior median differences and HPD intervals comparing the parameter values predicted by the MDHGLM among all predator experience levels. The parameter being compared is on the y axis, and the difference in parameter values between two experience levels is displayed on the x axis. The pairwise comparisons between experience levels are displayed across the three panels A, B, and C. The results are from the model that includes the outlier."}

knitr::include_graphics(
    here::here(
        "outputs", "04_outputs_figures",
        "appendix2_figureS1.png"
    )
)

```{r figS2, out.width = "100%", fig.cap = "**Figure S2.** Posterior median differences and HPD intervals comparing the trait correlations predicted by the MDHGLM between the advanced and novice levels. The difference is displayed on the y axis and the correlated parameters are displayed on the x axis. Positive values indicate that the correlation was greater when predators where advanced, while negative values indicate that the correlation was greater when predators where novices. The results are from the MDHGLM that includes the outlier."}

knitr::include_graphics(
  here::here(
    "outputs", "04_outputs_figures",
    "appendix2_figureS2.png"
  )
)
